Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(dplyr)
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup(age) %>%
  arrange(age)
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library(ggplot2)
ggplot(data = pf.fc_by_age_gender, aes(x=age, y=median_friend_count, color=gender)) + geom_line()
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
library(ggplot2)
ggplot(data = pf.fc_by_age_gender.wide, aes(x=age, y=female/male)) + geom_line() + geom_hline(yintercept = 1)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure / 365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(data=subset(pf, !is.na(year_joined.bucket)), aes(x=age, y=friend_count)) + 
    geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(data=subset(pf, !is.na(year_joined.bucket)), aes(x=age, y=friend_count)) + 
    geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean) + 
    geom_line(stat='summary', fun.y=mean, linetype=2)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, pf$tenure > 0), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
ggplot(data = subset(pf, tenure > 0), aes(x=tenure, y=friendships_initiated/tenure)) + geom_line(aes(color = year_joined.bucket), stat='summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

#ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
#       data = subset(pf, tenure > 0)) +
#  geom_line(aes(color = year_joined.bucket),
#            stat = "summary",
#            fun.y = mean)

#ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
#       data = subset(pf, tenure > 0)) +
#  geom_line(aes(color = year_joined.bucket),
#            stat = "summary",
#            fun.y = mean)

#ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
#       data = subset(pf, tenure > 0)) +
#  geom_line(aes(color = year_joined.bucket),
#            stat = "summary",
#            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

