Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
ggplot(data=pf, aes(x=age, y=friend_count)) + geom_point()
```

***

#### What are some things that you notice right away?
Response:
some people whose age above 90s has friends more than 1000; young people below 30 tend to have more friends than other ages
***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_jitter(alpha=1/20) + xlim(13, 90)

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20) + 
  xlim(13, 90) + coord_trans(y="sqrt")

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
ggplot(data=pf, aes(x=age, y=friendships_initiated)) + 
  geom_point(alpha=1/10, position = position_jitter(h = 0)) + xlim(13, 90) + 
  coord_trans(y = 'sqrt')

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
library(dplyr)
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
head(pf.fc_by_age, 20)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(data = pf.fc_by_age, aes(x=age, y=friend_count_mean)) +
  geom_line() + xlab("Age") + ylab("Average Friend Count")

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000)) +
  geom_point(alpha=1/20, position = position_jitter(h=0), color = 'orange') + 
  geom_line(stat='summary', fun.y=mean) +
  geom_line(stat='summary', fun.y=quantile, fun.args = list(probs = .1), linetype=2, color='blue') +
  geom_line(stat='summary', fun.y=quantile, fun.args = list(probs = .5), color='blue') +
  geom_line(stat='summary', fun.y=quantile, fun.args = list(probs = .9), linetype=2, color='blue')

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
with(pf, cor.test(age, friend_count, method="pearson"))

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age<=70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(data=pf, aes(x=www_likes_received, y=likes_received)) + geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(data=pf, aes(x=www_likes_received, y=likes_received)) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, .95)) +
  ylim(0, quantile(pf$likes_received, .95)) +
  geom_smooth(method='lm', color='red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received))
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(data = Mitchell, aes(x=Month, y=Temp)) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Month, Temp))

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, aes(x=Month, y=Temp)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- pf$age + (1-pf$dob_month/12)
```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>% 
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friedn_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
```

Programming Assignment
```{r Programming Assignment}
p1 <- ggplot(data = subset(pf.fc_by_age, age < 71), aes(x=age, y=friend_count_mean)) +
  geom_line() + 
  xlab("Age") + 
  ylab("Average Friend Count") +
  geom_smooth()

p2 <- ggplot(data = subset(pf.fc_by_age_months, age_with_months < 71), aes(x=age_with_months, y=friend_count_mean)) +
  geom_line() + 
  xlab("Age With Months") + 
  ylab("Average Friend Count") +
  geom_smooth()

library(gridExtra)
grid.arrange(p2, p1, ncol=1)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

