Wine Exploration Data Analysis
========================================================

I choose the white wine dataset to perform exploration data analysis. This dataset contains quality and chemical compositions for 4898 white wine. 

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(car)
library(vcd)
library(RColorBrewer)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wineData <- read.csv('wineQualityWhites.csv', na.strings = c('', 'NA'))
```

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}
dim(wineData)
```

```{r echo=FALSE, Univariate_Plots}
str(wineData)
```

```{r echo=FALSE, Univariate_Plots}
summary(wineData)
```
The dataset contains 13 variables and almost 5000 observations, no NAs existed. Not all variables are helpful for analyzing. According to the specifications in wineQualityInfo.txt, I will create a new variable called bound.sulfur.dioxide defined as total.sulfur.dioxide minus free.sulfur.dioxide. 

I will create two categorical variables called sweet_level and quality_level. According to  [Wikipedia](https://en.wikipedia.org/wiki/Sweetness_of_wine#Residual_sugar), a wine can be classified as four levels: dry(up to 4 $g/dm^3$), medium dry(up to 12 $g/dm^3$), medium(up to 45 $g/dm^3$) and sweet(more than 45 $g/dm^3$) by sweetness; the quality ranges from 3 to 9 and 6 is the median, so I will classify the quality into three types: low(qualities of 3, 4 and 5), medium(quality of 6) and high(quanlity of 7, 8 and 9), this may do some help for analyzing.

```{r echo=FALSE, Univariate_Plots}
# create a new variable called bound.sulfur.dioxide according to specifications in wineQualityInfo.txt
wineData$bound.sulfur.dioxide <- wineData$total.sulfur.dioxide - wineData$free.sulfur.dioxide

# create a new variable called sugar_level by cut of residual sugar, convert it into ordered factor
wineData$sugar_level <- cut(wineData$residual.sugar, breaks = c(-Inf, 4, 12, 45, Inf), labels = c('dry', 'medium dry', 'medium', 'sweet'))
wineData$sugar_level <- factor(wineData$sugar_level, ordered=TRUE, levels=c('dry', 'medium dry', 'medium', 'sweet'))

# create a new variable called quality_level by quality ranges, convert it into ordered factor
wineData$quality_level <- ifelse(wineData$quality < 6, 'low', ifelse(wineData$quality == 6, 'medium', 'high'))
wineData$quality_level <- factor(wineData$quality_level, ordered=TRUE, levels=c('low', 'medium', 'high'))

# select and rearrange variables
wineData <- select(wineData, quality, quality_level, fixed.acidity, volatile.acidity, citric.acid, residual.sugar, sugar_level, chlorides, free.sulfur.dioxide, bound.sulfur.dioxide, sulphates, density, pH, alcohol)
```

And this is the new structure.
```{r echo=FALSE, Univariate_Plots}
str(wineData)
```

```{r echo=FALSE, Univariate_Plots}
# histogram of quality scores
ggplot(data=wineData, aes(x=quality_level)) + geom_bar()
```
```{r echo=FALSE, Univariate_Plots}
table(wineData$quality_level)
```
The data contains 1640 observations of low quality wines, 2198 observations of medium quality wines and 1060 high quality of wines.

```{r echo=FALSE, Univariate_Plots}
# histogram of fixed acidity
pOrigin <- ggplot(data=wineData, aes(x=fixed.acidity)) + geom_histogram()
pBinned <- ggplot(data=wineData, aes(x=fixed.acidity)) + geom_histogram(binwidth = 0.12)
grid.arrange(pOrigin, pBinned, ncol=1)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$fixed.acidity)
```
Most wines have fixed acidity between 6 and 7.5 $g/dm^3$. There is a peak around 7 $g/dm^3$. Mean 6.855 $g/dm^3$ and median 6.80 $g/dm^3$; the maximum is 14.2 $g/dm^3$ and the minimum is 3.8 $g/dm^3$.

```{r echo=FALSE, Univariate_Plots}
# histogram of volatile acidity
ggplot(data=wineData, aes(x=volatile.acidity)) + geom_histogram(binwidth = 0.01)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$volatile.acidity)
```
Most wines have volatile acidity between 0.15 and 0.31 $g/dm^3$ approximately. The mininum value is not much less than the maxinum. A high level of this value may lead to an unpleasant taste, I will analyze it with quality later on.

```{r echo=FALSE, Univariate_Plots}
# histogram of citric acidity
ggplot(data=wineData, aes(x=citric.acid)) + geom_histogram(binwidth = 0.01)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$citric.acid)
```
Most wines have a citric acidity between 0.25 and 0.5 $g/dm^3$. The minimum value is 0.0, which means containing no citric acid. Later on I will analyze the relationship between citric acid and quality.

```{r echo=FALSE, Univariate_Plots}
# histogram of residual sugar
pSugarHist <- ggplot(data=wineData, aes(x=residual.sugar)) + geom_histogram(binwidth = 0.1)
pSugarHistScaled <- ggplot(data=wineData, aes(x=residual.sugar)) + geom_histogram(binwidth = 0.01) + scale_x_log10()
grid.arrange(pSugarHist, pSugarHistScaled, ncol=1)
```

```{r echo=FALSE, Univariate_Plots}
# boxplot of residual sugar
ggplot(data=wineData, aes(x='residual sugar', y=residual.sugar)) + geom_boxplot()
```

```{r echo=FALSE, Univariate_Plots}
summary(wineData$residual.sugar)
```
The residual sugar contributes to the sweetness of wine. I create an original histogram and a log-tranformed one to better observe it. Most of the wines contains residual sugar between 1 $g/dm^3$ and 10 $g/dm^3$, but some outliers exist. The boxplot indicates that median is around 5 and outliers have residual sugar more than 20 $g/dm^3$, some even has more than 60 $g/dm^3$, that's too sweet.

```{r echo=FALSE, Univariate_Plots}
ggplot(data=wineData, aes(x=wineData$sugar_level)) + geom_bar()
```

```{r echo=FALSE, Univariate_Plots}
table(wineData$sugar_level)
```
Most of wines are dry or medium dry, there are 825 wines of medium, and only 1 wine is sweet. 

```{r echo=FALSE, Univariate_Plots}
# histogram of chlorides
ggplot(data=wineData, aes(x=chlorides)) + geom_histogram(binwidth = 0.001)
```
```{r echo=FALSE, Univariate_Plots}
# boxplot of chlorides
ggplot(data=wineData, aes(x='chlorides', y=chlorides)) + geom_boxplot()
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$chlorides)
```
Chlorides is the amount of salt in wine. Most wines have chlorides between 0.025 and 0.0755 $g/dm^3$. The boxplot indicates that there are many outliers existed, the maximum value is 0.346 $g/dm^3$, but the variation in chlorides is small. Will a too-salty wine tastes bad? I will exlore it later on.

```{r echo=FALSE, Univariate_Plots}
# histogram of free sulfur dioxide
ggplot(data=wineData, aes(x=free.sulfur.dioxide)) + geom_histogram(binwidth = 1)
```

```{r echo=FALSE, Univariate_Plots}
# histogram of bound sulfur dioxide
ggplot(data=wineData, aes(x=bound.sulfur.dioxide)) + geom_histogram(binwidth = 1)
```

```{r echo=FALSE, Univariate_Plots}
# histogram of sulphates
ggplot(data=wineData, aes(x=sulphates)) + geom_histogram(binwidth = 0.01)
```
```{r echo=FALSE, Univariate_Plots}
summary(select(wineData, free.sulfur.dioxide, bound.sulfur.dioxide, sulphates))
```
Sulfur dioxide can prevent microbial growth and oxidation of wine. But it is somewhat harmful so the amount has to be limited. Most wines have free sulfur dioxide below 50 $mg/dm^3$, and bound sulfur dioxide one between 50 and 150 $mg/dm^3$. The maximum values of free and bound sulfur dioxide is 289 and 331 respectively, maybe this is a factor that affects the quality. Sulphates is an additive that can keep sulfur dioxide levels, most wines have between 0.4 and 0.6 $g/dm^3$.

```{r echo=FALSE, Univariate_Plots}
# histogram of density
ggplot(data=wineData, aes(x=density)) + geom_histogram(binwidth = 0.0001)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$density)
```
Most wine has a density around 1.00 $g/cm^3$.

```{r echo=FALSE, Univariate_Plots}
# histogram of pH
ggplot(data=wineData, aes(x=pH)) + geom_histogram(binwidth = 0.01)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$pH)
```
All the wine is naturally acidic. And most of them are in pH 3.0-3.3; minimum 2.72, maximum 3.82, mean and median are almost the same. This may correlates to acidities above.

```{r echo=FALSE, Univariate_Plots}
# histograms of alcohol
ggplot(data=wineData, aes(x=alcohol)) + geom_histogram(binwidth = 0.1)
```
```{r echo=FALSE, Univariate_Plots}
# histograms of alcohol
ggplot(data=wineData, aes(x=alcohol)) + geom_histogram(binwidth = 0.01)
```
```{r echo=FALSE, Univariate_Plots}
summary(wineData$alcohol)
```
Wine is a kind of low-alcohol drink. Most of them have alcohol between 9% and 12%. The minimum is 8% and the maximum is 14.2%, mean and median are 10.4% and 10.51% respectively. I guess alcohol is one of the important factors that will affect the quality.

# Univariate Analysis

### What is the structure of your dataset?

There are 4898 wines with 14 variables (quality, quality level, fixed acidity, volatile acidity, citric acid, residual sugar, sugar level, chlorides, free and total sulfur dioxide, sulphates, density, pH and alcohol).

quality level: low < medium < high
sugar level: dry < medium dry < medium < sweet

Other observations:

* Most wines are of medium quality level;
* The median quality is 6;
* Most wines have fixed acidity between 6 and 7.5 $g/dm^3$;
* Most wines have volatile acidity between 0.15 and 0.31 $g/dm^3$;
* Most wines have citric acid between 0.25 and 0.5 $g/dm^3$;
* Most wines are dry or medium dry;
* Most wines have chlorides between 0.025 and 0.0755 $g/dm^3$, but many outliers existed;
* Most wines have free sulfur dioxide below 50 $mg/dm^3$ and bound sulfur dioxide 50-150 $mg/dm^3$, a 0.4-0.6 $g/dm^3$ sulphates keep their levels;
* Most wines are naturally acidic, around 3.0-3.3;
* Wines are low-alcohol drinks, about 9%-12%.

### What is/are the main feature(s) of interest in your dataset?
Quality and quality level are the main features of interest in my dataset. I want to figure out how these chemical compositions affect the quality of wine. 

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Fixed acidity, volatile acidity, citric acid, residual sugar, sugar level, chlorides, sulfur dioxide, sulphates and alcohol are likely to contribute to quality of wines. 

### Did you create any new variables from existing variables in the dataset?
Yes, I create a new variable "bound.sulfur.dioxide" to divide total sulfur dioxide into two parts: the free one and the bound one, thus investigate them apartly in following explorations. And I create two categorical variables called "quality_level" and "sugar_level", the former is created by classifying qualities into three groups: low, medium and high; the latter is created by cutting residual sugar into four levels: dry, medium dry, medium and sweet.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
Yes, the histogram of residual sugar is highly right-skewed, this makes it difficult to analyze. So I create a log-transformed histogram to better investigate it. The transformed distribution appears bimodal with the residual sugar peaking at both 3 and 9 $g/dm^3$.

# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
cor(select(wineData, -quality_level, -sugar_level))
```

```{r echo=FALSE, warning=FALSE, Bivariate_Plots, fig.width=10}
scatterplotMatrix(~ quality + fixed.acidity + residual.sugar + chlorides + free.sulfur.dioxide + bound.sulfur.dioxide + density + pH + alcohol, data = wineData, spread = FALSE, smoother.args=list(lty=2), main='Scatter plot matrix of wine')
```
Some variables correlate with each other. Density correlates strongly with residual sugar and alcohol. All meaningful correlations are:

* quality: density(-0.307), alcohol(0.436)
* fixed acidity: pH(-0.43)
* residual sugar: free sulfur dioxide(0.30), bound sulfur.dioxide(0.34), density(0.84), alcohol(-0.45)
* chlorides: alcohol(-0.36)
* density: bound sulfur dioxide(0.50), alcohol(-0.78)
* alcohol: bound sulfur dioxide(-0.43)

Density and alcohol correlate with quality. I'd like to take a look at scatter plot involving them first.
```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=density, y=quality)) + geom_point()
```
```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=density, y=quality)) + 
  geom_jitter(alpha=0.02) + 
  scale_x_continuous(limits = c(0.95, 1.05)) + 
  geom_smooth(color='red', se=FALSE, method='lm')
```
The horizontal lines indicate quality values are all integers. After adding jitter, transparency, lineary smooth and changing the y limits, it shows a small and negative correlation between density and quality.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=density)) + geom_boxplot() + xlab('quality levels')
```

```{r echo=FALSE, Bivariate_Plots}
by(wineData$density, wineData$quality_level, summary)
```
The density median decreases as the quality level improves, and the variations are all very small. Wines of high quality level tends to have more outliers than the others. Density of wines is very close to 1.00 $g/cm^3$.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=alcohol, y=quality)) + geom_point()
```
```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=alcohol, y=quality)) + 
  geom_jitter(alpha=0.05) + 
  geom_smooth(se=FALSE, method='lm', color='red')
```
Similarly, After adding jitter, transparency and linear smooth, it shows a small and positive correlation between alcohol and quality. Next I'll investigate alcohol across different quality levels.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=alcohol, y=density)) + 
  geom_jitter(alpha=0.05) + 
  geom_smooth(se=FALSE, method='lm', color='red')
```
Alcohol correlates with density strongly, and the plot looks quite linear.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=alcohol)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$alcohol, wineData$quality_level, summary)
```
Low level wines have many outliers. And the alcohol median increases as the quality level improves. Wines of medium level have the biggest variation in alcohol and wines of low level have the smallest.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=fixed.acidity, y=pH)) + geom_point()
```
```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=fixed.acidity, y=pH)) + geom_jitter(alpha=0.05) + geom_smooth(method='lm', color='red', se=FALSE)
```
It is not surprising that fixed acidity correlates with pH, and after dealing with overplotting by adding jitter and alpha, the plot shows a meaningful-but-small correlation between them. The higher the fixed acidity, the lower the pH (more acid). So again let's take a look at fixed acidity across quality levels.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=fixed.acidity)) + geom_boxplot() + xlab('quality levels')
```

```{r echo=FALSE, Bivariate_Plots}
by(wineData$fixed.acidity, wineData$quality_level, summary)
```
The higher the quality level, the less the outliers. Both the median and variation remain approximately the same across quality levels.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=volatile.acidity)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$volatile.acidity, wineData$quality_level, summary)
```
A too high of levels of volatile acidity could lead to unpleasant taste, and this is proven by the outliers in plot. More outliers appear in low quality. The variation doesn't change very much, but wines of medium and high quality tends to have median lower than those of low quality.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=citric.acid)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$citric.acid, wineData$quality_level, summary)
```
Citric acid is yet another factor that may affect taste, it add freshness and falvor to wines. The median doesn't change very much as the quality level improves, but the variation becomes smaller and smaller, all three quality levels have outliers.

So after analyzing three different acidities, what is pH value like across quality levels?

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=pH)) + geom_boxplot() + xlab('quality levels')
```

```{r echo=FALSE, Bivariate_Plots}
by(wineData$pH, wineData$quality_level, summary)
```
The median and variation in pH increases slightly as the quality level improves. High quality level tends to have less outliers than others.

Although they show no correlation with quality, but I still want to investigate what sulphates, free and bound sulfur dioxide look like across quality levels. 

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=free.sulfur.dioxide)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$free.sulfur.dioxide, wineData$quality_level, summary)
```
As the quality level improves, the median decreases slightly and the variation decreases significantly, all three levels have outliers.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=bound.sulfur.dioxide)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$bound.sulfur.dioxide, wineData$quality_level, summary)
```
The median decreases as the quality level improves. And wines of high quality level have the smallest variation in bound sulfur dioxide. All three levels have outliers.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=sulphates)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$sulphates, wineData$quality_level, summary)
```
The median remains approximately the same. The variation in sulphates increases as the quality level improves. All three levels have outliers.

Next I'll consider about residual sugar. It is the amount of sugar remaining after fermentaion stops. At the very beginning I thought it might correlate with alcohol strongly but what is surprising is that it correlates with density most strongly.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=subset(wineData, residual.sugar <= quantile(wineData$residual.sugar, probs = .99)), aes(x=residual.sugar, y=density)) + geom_jitter(alpha=0.05) + geom_smooth(method='lm', color='red')
```
The relationship looks quite linear. Now it's time to analyze residual sugar across quality levels.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=residual.sugar)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$residual.sugar, wineData$quality_level, summary)
```
The median decreases significantly as the quality level improves, and the variation decreases too, indicating that the higher the quality levels, the less the residual sugar remains. Many advanced wines are dry, thus this is reasonable. Now let's take a closer look at relationships between quality levels and sugar levels, see if we can find something interesting.

```{r echo=FALSE, Bivariate_Plots}
mosaic(~sugar_level + quality_level, data=wineData, shade=TRUE, legend=TRUE)
```
From the mosaic plot we can confirm that quality levels do correlate with sugar levels. Wines of high quality tends to be dry.

```{r echo=FALSE, Bivariate_Plots}
ggplot(data=wineData, aes(x=quality_level, y=chlorides)) + geom_boxplot() + xlab('quality levels')
```
```{r echo=FALSE, Bivariate_Plots}
by(wineData$chlorides, wineData$quality_level, summary)
```
The variation in chlorides remains the same, but the median decreases as the quality level improves, all three levels have outliers but high quality level is less than the others.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

Density correlates with residual sugar and alcohol strongly, the former is positive, and the latter is negative.

Quality has a meaningful but small correlation with density and alcohol respectively. Density of all wines is close to 1.00 $g/cm^3$. 

Sugar level correlates with quality level, wines of high quality tends to be dry.

Other observations: 

* Wines of high quality tends to have smallest median and variation in density;
* Wines of high quality tends to have biggest median in alcohol;
* Wines of high quality tends to have least outliers in fixed acidity;
* Wines of high quality tends to have smallest median and least outliers in volatile acidity;
* Wines of high quality tends to have smallest median and variation in citric acid, but many outliers existed;
* Wines of high quality tends to have biggest median, variation and least outliers in pH;
* Residual sugar of wines decreases as the quality level improves;
* Chlorides of wines decreases slightly as the quality level improves;

Fixed acidity has a negative, meaningful but small correlation with pH, about -0.43.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

Fixed acidity tends to correlate with pH value. The higher the fixed acidity, the lower the pH value.

### What was the strongest relationship you found?

The residual sugar is positively and strongly correlated with density; alcohol is negatively correlates with density but a little bit less strongly. 

And if we want a better linear regression model between residual sugar and density, we must take bound sulfur dioxide and alcohol into consideration. Based on the $R^2$ value, if incorporated with bound sulfur dioxide and alcohol, residual sugar explains abount 91.4% of the variance in density.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(data=wineData, aes(x=quality, color=sugar_level)) + geom_density() + 
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'sugar level', reverse = T,
    override.aes = list(alpha = 1, size = 2)))
```
This density plot indicates that at the lower part of quality comes with more wines of sweet and medium, and higher part of quality comes with more wines of medium dry and dry. The only exception is at 4, where there are more dry and medium dry, so I think residual sugar is not the only factor that affects taste.

```{r echo=FALSE, Multivariate_Plots}
ggplot(data=wineData, aes(x=residual.sugar, y=citric.acid)) + geom_jitter(alpha=0.05) + facet_wrap(~quality_level, nrow=3)
```
Comparing with other levels, wines of high quality level tends to have conservative residual sugar and citric acid, the corresponding plot looks more centric: residual sugar is limited within 20 $g/dm^3$ and most citric acid within 0.5 $g/dm^3$.

```{r echo=FALSE, Multivariate_Plots}
ggplot(data=subset(wineData, residual.sugar <= quantile(wineData$residual.sugar, prob=.99)), 
       aes(x=residual.sugar, y=density)) + 
  geom_point() +
  geom_smooth(se=FALSE, method='lm', color='red') +
  facet_wrap(~quality_level)
```
```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density~residual.sugar, data=wineData))
```
What we can infer from the $R^2$ value is that residual sugar only explains about 70.4% of the variance in density. What about situations in different quality levels?

```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density~residual.sugar, data=subset(wineData, quality_level == 'low')))
```
For wines of low quality level, the residual sugar explains about 77.38% of the variance in density, this is a little improvement.

```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density~residual.sugar, data=subset(wineData, quality_level == 'medium')))
```
For wines of medium quality level, the residual sugar explains 73.21% of the variance in density, this is also a little improvement.

```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density~residual.sugar, data=subset(wineData, quality_level == 'high')))
```
For wines of high quality level, the model becomes worse, now it's only 67.27%. 

Density also correlates with alcohol strongly, next I'd like to analyze the same way as residual sugar.
```{r echo=FALSE, Multivariate_Plots}
ggplot(data=subset(wineData, alcohol <= quantile(wineData$alcohol, prob=.99)), 
       aes(x=alcohol, y=density)) + 
  geom_point() +
  geom_smooth(se=FALSE, method='lm', color='red') +
  facet_wrap(~quality_level)
```
```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density ~ alcohol, data=wineData))
```
In general, alcohol explains about 60.86% of the variance in density. Will it be better under different quality levels?

```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density ~ alcohol, data=subset(wineData, quality_level == 'low')))
```
```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density ~ alcohol, data=subset(wineData, quality_level == 'medium')))
```
For wines of low and medium quality, the model becomes worse, it can only explains about 46.26% and 55.28% of the variance in density respectively.

```{r echo=FALSE, Multivariate_Plots}
summary(lm(formula=density ~ alcohol, data=subset(wineData, quality_level == 'high')))
```
However, this model is better for wines of high quality, now $R^2$ improves to 0.7117, meaning that the model can explain about 71.17% of the variance in density.

I am not very content with these values. Trying to create models about density by just residual sugar or alcohol alone is not ideal. There must be a comprehensive model including more variables. By checking the correlation matrix again, I finally realize that there is a complex correlation existed for a variety of variables:

1. quality correlates with density and alcohol meaningfully but small;
2. density correlates with residual sugar and alcohol strongly;
3. density correlates with bound sulphur dioxide moderately;

So is it possible to create a better model by incorporating them all? Let's try this!
```{r echo=FALSE, Multivariate_Plots}
wineData$bound.sulfur.level <- cut(wineData$bound.sulfur.dioxide, breaks=c(4, 100, 331))
ggplot(data=subset(wineData, residual.sugar <= quantile(wineData$residual.sugar, prob=.99)), 
       aes(x=residual.sugar, y=density, color=alcohol, shape=bound.sulfur.level)) + 
  geom_point() + facet_wrap(~quality_level, nrow=3)
```

```{r echo=FALSE, Multivariate_Plots}
m1 <- lm(formula=density ~ residual.sugar, data=wineData)
m2 <- update(m1, ~ . + bound.sulfur.dioxide)
m3 <- update(m2, ~ . + alcohol)
summary(m3)
```
And yes! The $R^2$ improves to 0.9145, which means that residual sugar, together with bound sulfur dioxide and alcohol, explains 91.4% of the variance in density. It's much better now. Let's again take a look at situations of different quality levels.

```{r echo=FALSE, Multivariate_Plots}
m1 <- lm(formula=density ~ residual.sugar, data=subset(wineData, quality_level == 'low'))
m2 <- update(m1, ~ . + bound.sulfur.dioxide)
m3 <- update(m2, ~ . + alcohol)
summary(m3)
```
```{r echo=FALSE, Multivariate_Plots}
m1 <- lm(formula=density ~ residual.sugar, data=subset(wineData, quality_level == 'medium'))
m2 <- update(m1, ~ . + bound.sulfur.dioxide)
m3 <- update(m2, ~ . + alcohol)
summary(m3)
```
For wines of low and medium quality, this model become less effective, but can still explain about 89% of the variance in density.

```{r echo=FALSE, Multivariate_Plots}
m1 <- lm(formula=density ~ residual.sugar, data=subset(wineData, quality_level == 'high'))
m2 <- update(m1, ~ . + bound.sulfur.dioxide)
m3 <- update(m2, ~ . + alcohol)
summary(m3)
```
For wines of high quality, it improves again! Now this model can explain about 94.03% of the variance in density. These features strengthened each other in terms of this quality level.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

Although correlated strongly, the model is not very ideal if creating with density and only one feature(residual sugar or alcohol) alone, it will be even worse under some of the quality levels. By checking features that correlates with each other strongly or moderately, I finally create a new model that is much better. Especially under the high quality level, density, residual sugar, alcohol and bound sulfur dioxide strengthened each other.

### Were there any interesting or surprising interactions between features?
Yes, residual sugar and citric acid tend to be more conservative in wines of high quality, this is reasonable for residual sugar because sugar level correlates with quality level. What is surprising is that although citric acid can add freshness and flavor to wines, it is not "the more the better".

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
Yes, after doing some explorations and looking at relationships between different features. I create a model of density by taking residual sugar, alcohol and bound sulphur dioxide into account. It explains 91.4% of the variance in density in general. This model has limitations, it explains better under high quality level(about 94.03%), in levels of low and medium, it becomes worse, but can still explain about 89%.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data=wineData, aes(x=wineData$sugar_level)) + 
  geom_bar() + 
  labs(title='Wines of different sugar levels', x='Sugar level', y='Number of wines')
```

### Description One
Most wines are of dry and medium dry, there are more than 750 wines of medium, and wines of sweet are very rare.

### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
